Document Expansion with External Collections for Information Retrieval

### 2016-08-29 paper.tex
changed to article format
changed lambda to alpha for RM3
added background on RMs

\documentclass{report}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{caption}
\usepackage{enumitem}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{verbatim}
\usepackage{colortbl}
\usepackage[table]{xcolor}

@@

@@

@@

\begin{table}[htbp]
\centering
\begin{tabular}{|c|l|c|c|} \hline
{\bf Collection} & {\bf Run} & {\bf MAP} & {\bf nDCG@20} \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{AP} & {\it baseline-ql} & 0.2337 & 0.4170 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.3298 ($\lambda=0.2$) & 0.4852 ($\lambda=0.3$) \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & 0.3291 & 0.4858 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & ?? & ?? \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{Robust} & {\it baseline-ql} & 0.2185 & 0.3825 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.2349 ($\lambda=0.2$) & 0.3893 ($\lambda=0.6$) \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & 0.2663 & 0.3940 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & 0.2420$^{\Downarrow}$ & 0.4066$^{\uparrow}$ \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{wt10g} & {\it baseline-ql} & 0.1687 & 0.2753 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.1735 ($\lambda=0.7,0.8$) & 0.2738 ($\lambda=1.0$)\\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & 0.1683 & 0.2633 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & 0.1827 & 0.3153$^{\uparrow\Uparrow}$ \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{blogs06} & {\it baseline-ql} & 0.3064 & 0.2415 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.3103 ($\lambda=0.7$) & 0.2431 ($\lambda=1.0$)\\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & 0.3064 & 0.2330 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & 0.2988 & 0.2365 \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{clueweb09} & {\it baseline-ql} & 0.1117 & 0.1944 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.1163 ($\lambda=0.2$) & 0.2034 ($\lambda=0.7$) \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & ?? & ?? \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & 0.1183$^{\uparrow}$ & 0.2100 \\ \hline
\end{tabular}
\caption{The top-scoring runs and baselines by MAP. $\uparrow$ indicates statistically significant improvements over the oracle RM3 baseline, \textit{baseline-rm}, while $\Uparrow$ and $\Downarrow$ indicate statistically significant improvement and decrease compared to the cross validated RM3 baseline, \textit{baseline-rm-cv}. Significance at $p < 0.05$.}
\label{table.performance}
\end{table}

Retrieval performance of the baselines and top-scoring runs are shown in Table \ref{table.performance}. Mean average precision (MAP) and normalized discounted cumulative gain at 20 (nDCG@20) scores marked with $\uparrow$ are greater than the optimal RM3 run with statistical significance at $p < 0.05$ using a paired one-tailed t-test. We also run a more realistic RM3 run using 10-fold cross validation to set the mixing parameter $\lambda$. Scores marked with $\Uparrow and \Downarrow$ signal statistically significant increases and decreases in performance relative to this cross validated RM3 run. Note that baselines correspond to $\sum_{j=1}^n \lambda_{{\mathcal{E}_D}_j} = 0.0$.

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{caption}
\usepackage{enumitem}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{verbatim}
\usepackage{colortbl}
\usepackage[table]{xcolor}

@@

\subsubsection{Query Likelihood}

@@

Relevance modeling is a form of pseudo-relevance feedback that uses top ranked documents to estimate a language model representing documents relevant to a query \cite{Lavrenko2001}. This language model, known as a relevance model, acts as a form of query expansion and is generally used with the KL divergence retrieval model \cite{Zhai2006} to score documents.

A relevance model takes the form of

\begin{flalign}\label{equation.rm1}
	P(w|R) = \sum_{D \in C} P(D) P(w|D) P(Q|D)
\end{flalign}

\noindent where $P(Q|D)$ is calculated as in Equation \ref{equation.ql} and essentially weights word $w$ in document $D$ by the query likelihood of the document. Though theoretically calculated over all words and documents, relevance models are more efficient and robust when calculated over only the top terms in only the top ranked documents. These parameters are referred to as $fbTerms$ and $fbDocs$ respectively in Table \ref{table.parameters}.

Because relevance models are prone to query drift, research has shown that linearly interpolating a relevance model with the original query model to improve performance:

\begin{flalign}\label{equation.rm3}
	P(w|Q') = (1-\alpha) P(w|R) + \alpha P(w|Q).
\end{flalign}

\noindent $\alpha$ is a mixing parameter controlling the influence of the original query. This form of relevance model is known as ``RM3'' and is the baseline used throughout this paper.

@@

\begin{table}[htbp]
\centering
\begin{tabular}{|c|l|c|c|} \hline
{\bf Collection} & {\bf Run} & {\bf MAP} & {\bf nDCG@20} \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{AP} & {\it baseline-ql} & 0.2337 & 0.4170 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.3298 ($\alpha=0.2$) & 0.4852 ($\alpha=0.3$) \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & 0.3291 & 0.4858 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & ?? & ?? \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{Robust} & {\it baseline-ql} & 0.2185 & 0.3825 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.2349 ($\alpha=0.2$) & 0.3893 ($\alpha=0.6$) \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & 0.2663 & 0.3940 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & 0.2420$^{\Downarrow}$ & 0.4066$^{\uparrow}$ \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{wt10g} & {\it baseline-ql} & 0.1687 & 0.2753 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.1735 ($\alpha=0.7,0.8$) & 0.2738 ($\alpha=1.0$)\\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & 0.1683 & 0.2633 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & 0.1827 & 0.3153$^{\uparrow\Uparrow}$ \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{blogs06} & {\it baseline-ql} & 0.3064 & 0.2415 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.3103 ($\alpha=0.7$) & 0.2431 ($\alpha=1.0$)\\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & 0.3064 & 0.2330 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & 0.2988 & 0.2365 \\ \hline\hline
\rule{0pt}{2.5ex} \multirow{3}{*}{clueweb09} & {\it baseline-ql} & 0.1117 & 0.1944 \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm} & 0.1163 ($\alpha=0.2$) & 0.2034 ($\alpha=0.7$) \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it baseline-rm-cv} & ?? & ?? \\ \cline{2-4}
\rule{0pt}{2.5ex} & {\it exp-ql} & 0.1183$^{\uparrow}$ & 0.2100 \\ \hline
\end{tabular}
\caption{The top-scoring runs and baselines by MAP. $\uparrow$ indicates statistically significant improvements over the oracle RM3 baseline, \textit{baseline-rm}, while $\Uparrow$ and $\Downarrow$ indicate statistically significant improvement and decrease compared to the cross validated RM3 baseline, \textit{baseline-rm-cv}. Significance at $p < 0.05$.}
\label{table.performance}
\end{table}

Retrieval performance of the baselines and top-scoring runs are shown in Table \ref{table.performance}. Mean average precision (MAP) and normalized discounted cumulative gain at 20 (nDCG@20) scores marked with $\uparrow$ are greater than the optimal RM3 run with statistical significance at $p < 0.05$ using a paired one-tailed t-test. We also run a more realistic RM3 run using 10-fold cross validation to set the mixing parameter $\alpha$. Scores marked with $\Uparrow and \Downarrow$ signal statistically significant increases and decreases in performance relative to this cross validated RM3 run. Note that baselines correspond to $\sum_{j=1}^n \lambda_{{\mathcal{E}_D}_j} = 0.0$.
